---
title: "Fitting Poisson Model to Epilepsy Data"
output: rmarkdown::html_vignette
description: >
  This vignette describes how we fits the poisson model to epilepsy data with a funtion 
  that provides summary information including the point estimates, standard erros, 
  and 95% confidence intervals for each parameter in the model.
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(FinalProjectComputing)
library(ggplot2)
library(tidyverse)
library(lme4)
```



 # Exploratory data analysis
 
```{r include=FALSE}
data("epilepsy")
epilepsy <- epilepsy %>%
group_by(id,treat,expind,age) %>%
summarize(seizures = sum(seizures),.groups = "drop")
```


summary statistics
```{r echo=FALSE}
summary(epilepsy)
```
```{r echo=FALSE}
str(epilepsy)
```
```{r echo=FALSE}
head(epilepsy)
```
There are 4 variables which are treat, expind, age, and seizures with 59 subjects ID

age - age of the patients as a continuous predictor
expind - a categorical variable with two levels (0 for before and 1 for after)
expind :treat - a categorical variable with two levels (1 for observations from individuals on the drug in the after period and 0 otherwise)
seizures - the number of seizures patients had as continuous variable

```{r echo=FALSE}
ggplot(epilepsy , aes(x= treat , y= seizures)) +
geom_boxplot(aes(group=treat))
```
```{r echo=FALSE}
ggplot(epilepsy , aes(x= expind , y= seizures)) +
geom_boxplot(aes(group=expind))
```
Two box-plots show the number of seizures between two groups of treatment and 
between two groups of expind.

```{r}
plot(epilepsy$age, epilepsy$seizures,
        xlab="age",
        ylab="seizures")
```
Age and seizures seem there is no linear relationship. 


# Model Fitting and use of the functions

The purpose of this package is to use the epilepsy data to fit the following model

\begin{equation*}
    Y_{ij} \sim \mathrm{Poisson}(\mathrm{exp}(\mu_{ij})) 
\end{equation*}

With $Y_{ij}$ being the number of seizures of an individual and $\mu_{ij}$ following this equation

$$
    \eta_{ij} =\mathrm{log}(\mu_{ij}) = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{3j} + \beta_3 x_{2i}x_{3j} + z_i
$$

where 
\begin{align*}
    x_{i1} & = \text{ Age of patient} \\
    x_{2i} & = \text{ 0 if placebo, 1 otherwise} \\
    x_{3j} & = \text{0 before treatment, 1 after} \\
    z_i & = \text{Random effect with mean 0 and unknown variance}
\end{align*}

To fit the model and get the main results we can use the function run_model specifying that the epilepsy data.

```{r}
run_model(epilepsy,"epilepsy")
```

To get the estimates of the standard error and confidence intervals using bootstraping we can use the function model_results. Note that this functions need the parameter n_boot that is the number of bootstraps to do.

```{r warning=FALSE}
model_results(epilepsy,300)
```

Finally, to get the p-values of the following hypothesis testing for each predictor

$$
H_0: \beta_k = 0 \text{ vs } H_1: \beta_k \neq 0
$$

we use the function model_summary

```{r warning=FALSE}
model_summary(epilepsy,300)
```

It is also possible to get the p-Value of only one of the predictors using the function pValue_predictor replacing "predictor" by the name. In the case of the predictor age with the t statistic 1.9 we can use it like this

```{r warning=FALSE}
pValue_age(epilepsy,1.9,300)
```


